---
title: "multivariate-benchmarking"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{multivariate-benchmarking}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(imputeTestbench)
```

## Introduction

The `imputeTestbench` package provides tools to compare imputation methods for time series data. 
Originally, the `impute_errors()` function works on a single time series. 
This vignette introduces `mv_impute_errors()`, a multivariate wrapper that applies 
the same benchmarking to each column of a matrix or data frame, returning a named list 
of results. This makes it easy to compare imputation performance across multiple 
related time series simultaneously.

## Example with stock market data

We'll use the built-in `EuStockMarkets` dataset, which contains daily closing prices of four European stock indices: DAX, SMI, CAC, and FTSE.

```{r load}
library(imputeTestbench)
data("EuStockMarkets")
# Preview the data
head(EuStockMarkets)
```

## Running the multivariate benchmark

We call `mv_impute_errors()` with a small range of missing percentages and few repetitions to keep the vignette fast.

```{r run}
mv_results <- mv_impute_errors(
  multi_dataIn = EuStockMarkets,
  missPercentFrom = 10,
  missPercentTo = 30,
  interval = 10,
  repetition = 3   # low for vignette speed
)
```


## Exploring the output

The output is a named list, where each element corresponds to a column of the input. Each element is itself an `errprof` object, identical to what `impute_errors()` returns for a univariate series.

```{r explore}
# Names of the list (the stock indices)
names(mv_results)

# Results for the DAX index
mv_results$DAX

# Average RMSE for the 'na.approx' method on DAX
mv_results$DAX$na.approx

# Access the raw errors (all repetitions) for SMI
str(attr(mv_results$SMI, "errall"), max.level = 2)
```

## Comparing across variables

We can extract the average errors for all variables into a single data frame for easy comparison.

```{r compare}
compare_df <- do.call(rbind, lapply(names(mv_results), function(var) {
  res_var <- mv_results[[var]]
  data.frame(
    variable = var,
    missPercent = res_var$MissingPercent,
    approx = res_var$na.approx,
    interp = res_var$na.interp,
    locf = res_var$na.locf,
    mean = res_var$na_mean
  )
}))
head(compare_df)
```
